---
title: "PA1_template.Rmd"
author: "Jiangpeng Pan"
date: "6/14/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Loading and preprocessing the data
Show any code that is needed to

1. Load the data (i.e.read.csv())
Process/transform the data (if necessary) into a format suitable for your analysis


```{r}
if (!file.exists("activity.csv") )
    {
     url <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(url,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }

data <- read.csv("activity.csv") 
```

What is mean total number of steps taken per day?
  1. Calculate the total number of steps taken per day
  2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
  3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
StepsByDays <- aggregate(steps ~ date, data, sum)
###```{r pressure, echo=FALSE}
hist(StepsByDays$steps, main = "Total Steps per day", xlab = "number of steps")

```
```{r}
### mean and median 
mean1 <- mean(StepsByDays$steps)
mean1
median1 <- median(StepsByDays$steps)
median1
```

the mean is `r mean1`, the median is `r median1` . 

What is the average daily activity pattern?
  1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
  2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
  
```{r}
interval <- aggregate(steps ~ interval, data, mean)
plot(interval$interval, interval$steps, type = 'l', main = "average number of steps by interval", xlab = 'Inverval', ylab = "number of steps")

```
```{r}
## maximum 
max_interval <- interval[which.max(interval$steps), 1]
max_interval
```
the interval with most steps is `r max_interval`. 

Imputing missing values
  1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
totalNA <- sum(!complete.cases(data))
totalNA
```
the total number of missing values are `totalNA`. 

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
step_mean <- aggregate(steps ~ interval, data, mean)
fill_NA <- numeric()
for (i in 1:nrow(data)) {
  obs <- data[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(step_mean, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fill_NA <- c(fill_NA, steps)
}
```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
new <- data
new$steps <- fill_NA
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
new_step <- aggregate(steps ~ date, new, sum, na.rm = TRUE)
hist(new_step$steps, main = "total steps", breaks = 15, xlab = 'number of steps')
``` 
```{r}
new_mean <- mean(new_step$steps)
new_median <- median(new_step$steps)
new_mean_diff <- new_mean - mean1
new_median_diff <- new_median - median1
```
the new mean is `r new_mean`, and the new median is `r new_median`. the difference of mean is `r new_mean_diff`, and the difference of medain is `r new_median_diff`. 

Are there differences in activity patterns between weekdays and weekends?

  1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
library(lattice)
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
new$dow = as.factor(ifelse(is.element(weekdays(as.Date(new$date)), weekdays), "Weekday", "Weekend"))
last_Total <- aggregate(steps ~ interval + dow, new, mean)
xyplot(last_Total$steps ~ last_Total$interval|last_Total$dow, main = "Average steps by interval", xlab = "interval", ylab = "number of steps", layout = c(1,2), type = "l")
```
